---
title: "Farm Household features and Data Base development"
author: "Francisco Fernandez"
date: "Friday, July 08, 2016"
output: html_document
subtitle: From Cluster Analysis to Farm-Household Model Data Base 
---

# Data Base construction
Based on PCA results and cluster analysis we developed the Final Data Base used to assess the economic impact of climate change on the agricultural sector of the Maule region of Chile at micro-level scale.

The data base development here presented is described in detail in this appendix.In a first place, joining the data survey and the results obtained from the cluster analysis we add a type number (which represent each household type) for each observation of RawData. After this, we select the main variables that we will use for the Data Base construction, where the information of each representative household will be presented.  

```{r,echo=FALSE}
### Read Data
FinalCluster<-read.table(file.choose("Final_cluster.csv"), header = TRUE, sep =",")

```


```{r,echo=FALSE,include=FALSE}
library(dplyr)
library(tidyr)

### Loading Data 
#FinalCluster<-read.table(file.choose("Final_cluster.csv"), header = TRUE, sep =",")
#RawData <-read.table(file.choose("DB_PMP_FINAL.csv"), header = TRUE, sep =";")

load("RawData.Rda")
#load("FinalCluster.Rda")
# Data frame folio and Typo
DF_FT  <- FinalCluster %>%
        select(folio, typo)

## Adding variables to RawData for Cluster Analysis
### Add Tot_Lab "Total Labor (working days per ha)"
RawData <- RawData %>%
        mutate(Tot_Lab = (jh_pt_f + jh_pt_c + jh_s_f + jh_s_c + jh_apf_f + jh_apf_c + jh_ahp_f + jh_ahp_c +
                       jh_r_f + jh_r_c + jh_c_f + jh_c_c)/x0)


### Add Hired_Lab "Hired Labor (working days per ha)"
RawData <- RawData %>%
        mutate(Hrd_Lab = (jh_pt_c +  jh_s_c +  jh_apf_c + jh_ahp_c + jh_r_c + jh_c_c)/x0)


### Add Hrd_ratio "Hired ratio (working days per ha) -> Hired labour/Total Labour"
RawData <- RawData %>%
        mutate(Hrd_ratio = Hrd_Lab / Tot_Lab )

### Add Family_Lab "Family Labor (working days per ha)"
RawData <- RawData %>%
        mutate(Fam_Lab = (jh_pt_f +  jh_s_f +  jh_apf_f + jh_ahp_f + jh_r_f + jh_c_f)/x0)

# Data frame Labour
DF_Lab <- RawData %>%
        select(folio, Fam_Lab, Hrd_Lab)

# Add "typo" variable to RawData
DF_FstClus  <- left_join(RawData, DF_FT, by = "folio")

#Select variables for Data Base
DF_FstClus2 <- DF_FstClus %>%
        select(typo, comuna, cultivo, sistema, x0, yield, p0, jh_c, jh, ctomaq, ctoinsumosh, ctototal)

#Add Labour Data to Cluster Data Frame
DF_FstClus3 <- bind_cols(DF_FstClus2, DF_Lab)
DF_FstClus3  <- DF_FstClus3 %>%
        select(typo, comuna, cultivo, sistema, x0, yield, p0, Fam_Lab, Hrd_Lab, jh_c, jh, ctomaq, ctoinsumosh, ctototal)

# Adding Labour cost ($CLP/ha)
DF_FstClus4  <- DF_FstClus3 %>%
        mutate(HrdLabCst = Hrd_Lab*jh_c, FamLabCst = Fam_Lab*jh) %>%
        mutate(TtlCst= HrdLabCst + FamLabCst + ctomaq + ctoinsumosh)

```


# Household type description
In the next tables we present the summary for each one of the cluster developed. 

```{r,echo=F,fig.cap="Table B1: Main characteristics of the farm types"}
DataStat <- data.frame(n= numeric(0), Farm_size= integer(0), mean_Rev = integer(0), Grn_shr = integer(0), SC_shr = integer(0), SV_shr = integer(0))
```


```{r,echo=FALSE}
Cluster1 <- FinalCluster %>%
        filter(typo == 1)

Cluster1$OC_shr<-NULL
Cluster1$SD_shr<-NULL

Cluster2 <- FinalCluster %>%
        filter(typo == 2)

Cluster2$OC_shr<-NULL
Cluster2$SD_shr<-NULL


Cluster3 <- FinalCluster %>%
        filter(typo == 3)

Cluster3$OC_shr<-NULL
Cluster3$SD_shr<-NULL


Cluster4 <- FinalCluster %>%
        filter(typo == 4)

Cluster4$OC_shr<-NULL
Cluster4$SD_shr<-NULL


#Cluster5 <- FinalCluster %>%
#        filter(typo == 5)

#Cluster5$OC_shr<-NULL
#Cluster5$SD_shr<-NULL


```

```{r,echo=F,fig.cap="tABLE b1:"}
# Number of farms 
DataStat[1,1] <- nrow(Cluster1)
DataStat[2,1] <- nrow(Cluster2)
DataStat[3,1] <- nrow(Cluster3)
DataStat[4,1] <- nrow(Cluster4)
#DataStat[5,1] <- nrow(Cluster5)

# Mean of farm size
DataStat[1,2] <- round(mean(Cluster1$Farm_size),2)
DataStat[2,2] <- round(mean(Cluster2$Farm_size),2)
DataStat[3,2] <- round(mean(Cluster3$Farm_size),2)
DataStat[4,2] <- round(mean(Cluster4$Farm_size),2)
#DataStat[5,2] <- round(mean(Cluster5$Farm_size),2)

DataStat[1,3] <- round(mean(Cluster1$Ttl_Rev),2)
DataStat[2,3] <- round(mean(Cluster2$Ttl_Rev),2)
DataStat[3,3] <- round(mean(Cluster3$Ttl_Rev),2)
DataStat[4,3] <- round(mean(Cluster4$Ttl_Rev),2)
#DataStat[5,3] <- round(mean(Cluster5$Ttl_Rev),2)

DataStat[1,4] <- round(mean(Cluster1$Grn_shr),2)
DataStat[2,4] <- round(mean(Cluster2$Grn_shr),2)
DataStat[3,4] <- round(mean(Cluster3$Grn_shr),2)
DataStat[4,4] <- round(mean(Cluster4$Grn_shr),2)
#DataStat[5,4] <- round(mean(Cluster5$Grn_shr),2)

DataStat[1,5] <- round(mean(Cluster1$SC_shr),2)
DataStat[2,5] <- round(mean(Cluster2$SC_shr),2)
DataStat[3,5] <- round(mean(Cluster3$SC_shr),2)
DataStat[4,5] <- round(mean(Cluster4$SC_shr),2)
#DataStat[5,5] <- round(mean(Cluster5$SC_shr),2)

DataStat[1,6] <- round(mean(Cluster1$SV_shr),2)
DataStat[2,6] <- round(mean(Cluster2$SV_shr),2)
DataStat[3,6] <- round(mean(Cluster3$SV_shr),2)
DataStat[4,6] <- round(mean(Cluster4$SV_shr),2)
#DataStat[5,6] <- round(mean(Cluster5$SV_shr),2)


DataStat

```


```{r,echo=FALSE}

Household1 <- DF_FstClus4 %>%
        filter(typo == 1)

Household1<-na.omit(Household1)

Household1 <- Household1 %>%
        group_by(comuna, cultivo, sistema) %>%
        summarise(area = sum(x0), yield = mean(yield), CropPrice = mean(p0), HiredLabor = mean(Hrd_Lab),
                  FamilyLab = mean(Fam_Lab), PriceHrdLab = mean(jh_c), PriceFamLab = mean(jh),
                  rentedMachCosts = mean(ctomaq), InputCosts = mean(ctoinsumosh))

CrpsHT1 <- Household1 %>%
        group_by(cultivo) %>%
        summarise(m_Area = sum(area)) %>%
        mutate(Area_shr = m_Area/sum(m_Area))

SysHT1  <- Household1 %>%
        group_by(sistema) %>%
        summarise(m_Sys = sum(area)) %>%
        mutate(Sys_shr = m_Sys/sum(m_Sys))

RevHT1  <- Household1 %>%
        group_by(cultivo) %>%
        summarise(m_Rev = sum(area * yield * CropPrice)) %>%
        mutate(Rev_shr = m_Rev/sum(m_Rev))

Household1$HT  <- "H1"

Household2 <- DF_FstClus4 %>%
        filter(typo == 2)

Household2<-na.omit(Household2)

Household2 <- Household2 %>%
        group_by(comuna, cultivo, sistema) %>%
        summarise(area = sum(x0), yield = mean(yield), CropPrice = mean(p0), HiredLabor = mean(Hrd_Lab),
                  FamilyLab = mean(Fam_Lab), PriceHrdLab = mean(jh_c), PriceFamLab = mean(jh),
                  rentedMachCosts = mean(ctomaq), InputCosts = mean(ctoinsumosh))

CrpsHT2 <- Household2 %>%
        group_by(cultivo) %>%
        summarise(m_Area = sum(area)) %>%
        mutate(Area_shr = m_Area/sum(m_Area))

SysHT2  <- Household2 %>%
        group_by(sistema) %>%
        summarise(m_Sys = sum(area)) %>%
        mutate(Sys_shr = m_Sys/sum(m_Sys))
                  
RevHT2  <- Household2 %>%
        group_by(cultivo) %>%
        summarise(m_Rev = sum(area * yield * CropPrice)) %>%
        mutate(Rev_shr = m_Rev/sum(m_Rev))

Household2$HT  <- "H2"


######################################--Household Type 3--#############################

Household3 <- DF_FstClus4 %>%
        filter(typo == 3)

Household3<-na.omit(Household3)

Household3 <- Household3 %>%
        group_by(comuna, cultivo, sistema) %>%
        summarise(area = sum(x0), yield = mean(yield), CropPrice = mean(p0), HiredLabor = mean(Hrd_Lab),
                  FamilyLab = mean(Fam_Lab), PriceHrdLab = mean(jh_c), PriceFamLab = mean(jh),
                  rentedMachCosts = mean(ctomaq), InputCosts = mean(ctoinsumosh))

CrpsHT3 <- Household3 %>%
        group_by(cultivo) %>%
        summarise(m_Area = sum(area)) %>%
        mutate(Area_shr = m_Area/sum(m_Area))

SysHT3  <- Household3 %>%
        group_by(sistema) %>%
        summarise(m_Sys = sum(area)) %>%
        mutate(Sys_shr = m_Sys/sum(m_Sys))
                  
RevHT3  <- Household3 %>%
        group_by(cultivo) %>%
        summarise(m_Rev = sum(area * yield * CropPrice)) %>%
        mutate(Rev_shr = m_Rev/sum(m_Rev))
                  

Household3$HT  <- "H3"

######################################--Household Type 4--#############################

Household4 <- DF_FstClus4 %>%
        filter(typo == 4)

Household4<-na.omit(Household4)

Household4 <- Household4 %>%
        group_by(comuna, cultivo, sistema) %>%
        summarise(area = sum(x0), yield = mean(yield), CropPrice = mean(p0), HiredLabor = mean(Hrd_Lab),
                  FamilyLab = mean(Fam_Lab), PriceHrdLab = mean(jh_c), PriceFamLab = mean(jh),
                  rentedMachCosts = mean(ctomaq), InputCosts = mean(ctoinsumosh))

CrpsHT4 <- Household4 %>%
        group_by(cultivo) %>%
        summarise(m_Area = sum(area)) %>%
        mutate(Area_shr = m_Area/sum(m_Area))

SysHT4  <- Household4 %>%
        group_by(sistema) %>%
        summarise(m_Sys = sum(area)) %>%
        mutate(Sys_shr = m_Sys/sum(m_Sys))
                  
RevHT4  <- Household4 %>%
        group_by(cultivo) %>%
        summarise(m_Rev = sum(area * yield * CropPrice)) %>%
        mutate(Rev_shr = m_Rev/sum(m_Rev))

Household4$HT  <- "H4"

######################################--Household Type 4--#############################

#Household5 <- DF_FstClus4 %>%
#        filter(typo == 5)

#Household5<-na.omit(Household5)

#Household5 <- Household5 %>%
#        group_by(comuna, cultivo, sistema) %>%
#        summarise(area = sum(x0), yield = mean(yield), CropPrice = mean(p0), HiredLabor = mean(Hrd_Lab),
#                  FamilyLab = mean(Fam_Lab), PriceHrdLab = mean(jh_c), PriceFamLab = mean(jh),
#                  rentedMachCosts = mean(ctomaq), InputCosts = mean(ctoinsumosh))

#CrpsHT5 <- Household5 %>%
#        group_by(cultivo) %>%
#        summarise(m_Area = sum(area)) %>%
#        mutate(Area_shr = m_Area/sum(m_Area))

#SysHT5  <- Household5 %>%
#        group_by(sistema) %>%
#        summarise(m_Sys = sum(area)) %>%
#                          mutate(Sys_shr = m_Sys/sum(m_Sys))
                  
#RevHT5  <- Household5 %>%
#        group_by(cultivo) %>%
#        summarise(m_Rev = sum(area * yield * CropPrice)) %>%
#        mutate(Rev_shr = m_Rev/sum(m_Rev))
                  
#Household5$HT  <- "H5"
                  
```

```{r}
CrpsHT1
SysHT1
RevHT1
```

```{r}
CrpsHT2
SysHT2
RevHT2
```

```{r}
CrpsHT3
SysHT3
RevHT3
```

```{r}
CrpsHT4
SysHT4
RevHT4
```

```{r}
#CrpsHT5
#SysHT5
#RevHT5
```

```{r}
################ Binding Households Data frames --> Creating Final DB for household model

FinalDB_0208 <- bind_rows(Household1, Household2, Household3, Household4)
FinalDB_0208 <- FinalDB_0208[,c(13,1:12)]
FinalDB_0208 <- FinalDB_0208 %>%
        rename(commune=comuna, crop=cultivo, system=sistema)


### Final Details
## Codes for communes
FinalDB_0208$commune <- as.character(FinalDB_0208$commune)
FinalDB_0208$commune[FinalDB_0208$commune == "Pencahue"] <- "PEN"
FinalDB_0208$commune[FinalDB_0208$commune == "Cauquenes"] <- "CAU"
FinalDB_0208$commune[FinalDB_0208$commune == "San Clemente"] <- "SC"
FinalDB_0208$commune[FinalDB_0208$commune == "Parral"] <- "PAR"

## Codes for systems
FinalDB_0208$system <- as.character(FinalDB_0208$system)
FinalDB_0208$system[FinalDB_0208$system == "riego"] <- "irr"
FinalDB_0208$system[FinalDB_0208$system == "secano"] <- "dry"


## Codes for Crops
FinalDB_0208$crop <- as.character(FinalDB_0208$crop)
FinalDB_0208$crop[FinalDB_0208$crop == "ARROZ"] <- "ric"
FinalDB_0208$crop[FinalDB_0208$crop == "ARVEJA"] <- "pea"
FinalDB_0208$crop[FinalDB_0208$crop == "AVENA"] <- "oat"
FinalDB_0208$crop[FinalDB_0208$crop == "CEBOLLA"] <- "oni"
FinalDB_0208$crop[FinalDB_0208$crop == "GARBANZO"] <- "chk"
FinalDB_0208$crop[FinalDB_0208$crop == "LENTEJA"] <- "len"
FinalDB_0208$crop[FinalDB_0208$crop == "MAIZ"] <- "mze"
FinalDB_0208$crop[FinalDB_0208$crop == "MELON"] <- "mel"
FinalDB_0208$crop[FinalDB_0208$crop == "PAPA"] <- "pot"
FinalDB_0208$crop[FinalDB_0208$crop == "POROTO"] <- "cmb"
FinalDB_0208$crop[FinalDB_0208$crop == "POROTO_VER"] <- "gbn"
FinalDB_0208$crop[FinalDB_0208$crop == "SANDIA"] <- "wtm"
FinalDB_0208$crop[FinalDB_0208$crop == "TOMATE"] <- "tom"
FinalDB_0208$crop[FinalDB_0208$crop == "TRIGO"] <- "wht"
FinalDB_0208$crop[FinalDB_0208$crop == "ZAPALLO"] <- "sqh"

## Rounding values
FinalDB_0208$yield <-round(FinalDB_0208$yield,2) 
FinalDB_0208$CropPrice <-round(FinalDB_0208$CropPrice)
FinalDB_0208$HiredLabor <- round(FinalDB_0208$HiredLabor,2)
FinalDB_0208$FamilyLab <- round(FinalDB_0208$FamilyLab,2)
FinalDB_0208$PriceHrdLab <- round(FinalDB_0208$PriceHrdLab)
FinalDB_0208$PriceFamLab <- round(FinalDB_0208$PriceFamLab)
FinalDB_0208$rentedMachCosts <- round(FinalDB_0208$rentedMachCosts)
FinalDB_0208$InputCosts <- round(FinalDB_0208$InputCosts)

tbl_df(FinalDB_0208)

#Saving FinalDB as csv
write.csv(FinalDB_0208, file="csv_files/FinalDB_0208.csv") 

```


